DP Problems of Interest

### 2012-12-01 docs/roadmap/design.tex
Some more docs.

@@

Let one parser reversal complexity be $O(p)$, single backtrack has $O(2n\cdot p)$ complexity. For the $k$-elements backtrack, since we regenerate all possible solutions, that is $O(k^{c+1})$ candidates (with $c$ the maximal number of concatenation in the parser), the overall complexity is $O(2n \cdot k^{c+1} \cdot p)$. Hence there is a $k^{c+1}$ factor to pay if we want to backtrack the $k$ best solutions\footnote{Note that the same factor lies in the matrix computation complexity.}.

%Algo: Read matrix, reapply the parser but by maintaining list of origins. Filter results to match multiplicities

>>>>

explain split into signature and grammar

@@

Let one parser reversal complexity be $O(p)$, single backtrack has $O(2n\cdot p)$ complexity. For the $k$-elements backtrack, since we regenerate all possible solutions, that is $O(k^{c+1})$ candidates (with $c$ the maximal number of concatenation in the parser), the overall complexity is $O(2n \cdot k^{c+1} \cdot p)$. Hence there is a $k^{c+1}$ factor to pay if we want to backtrack the $k$ best solutions\footnote{Note that the same $k^{c+1}$ factor lies in the forward matrix computation complexity.}.

\subsubsection{Backtrack utilization}
Since the dynamic programming input may be a subset of the larger problem that we want to resolve\footnote{For instance in matrix chain multiplication, we only care about matrix dimensions for dynamic programming, however, we ultimately want to multiply the real matrices and obtain a result.}, we need to be able to apply the result of the dynamic programming computation on values in a different domain. The easiest way to do that, is to reuse the same grammar on a different domain, and only compute the resulting trace obtained from the DP solver.

This step is pretty straightforward: since ADP parsers emphasize on the split between signature and grammar and decouples them, we only need to modify the signature to operate on another domain, and reuse the same grammar. The key point here is to notice that a backtrack trace on one parser can be reused on another, providing that they have the same grammar. For instance, to compute optimally a matrix chain multiplication, we solve the DP problem in a domain where matrices are represented by their dimensions, we obtain an optimal trace and feed it to a parser operating on the <<real matrices>> domain that will do the actual computation.

### 2012-12-01 docs/roadmap/problems.tex
Algebrae systematic product.
Some more docs.

% ----------------------------------------------
\newpage
\subsubsection{Convex polygon triangulation}\ol
\item Problem: triangulating a polygon of $n$ vertices with least total cost for added edges. We denote the cost of adding an edge between the  pair of edges $i,j$ by $S(i,j)$, Where $S_{n \times n}$ is a lower triangular matrix compacted in memory (rows are contiguous) with a 0 diagonal that is omitted \footnote{Arbitrary convention for both architectural implementation and code generator. Rationale: in lower triangular matrix, element address is independent of the matrix size.}, hence $|S|=\tfrac{n^2}{2}=N$.

\item Matrices: $M_{n\times 2n}, B_{n \times 2n}$ \textit{<<first edge, last edge>>} upper triangular including main diagonal 
\item Alphabets:\ul
	\item Input: $\Sigma(S_{(i,j)})=\{0..m\}$ with $m=\max_S(i,j) \forall i,j$ determined at runtime\footnote{We need to scan/have stats about $S$ and that's where LMS plays a role}.
	\item Cost matrix: $\Sigma(M)=\{0..z\}$ with $z = m \cdot (n-2)$ (we add at most $n-2$ edges).
	\item Backtrack matrix: $\Sigma(B)=\{stop, 0..n\}$ (the index of the edge we add)
	\ule
\item Initialization: $M_{(i,i)}=0, M_{(i,i+1)}=0, B_{(i,i)}=stop \quad\forall i$

\item Recurrence: \[M_{(i,j)}=\left\{ S(i,j) + \max_{i<k<j}M_{(i,k)}+M_{(k,j)} \,\,\rule[-.75em]{.5pt}{2em}\,\,  k \right\} = B_{(i,j)} \]
	It is interesting to note that even in the sequential world, this problem is solved 
	by filling the diagonals, ie. computing sub-solutions for all polygons of size $k$ before
	those of size $k+1$.

@@

\item Visualization: the layout is the a matrix of size $n \times (2n-2)$, because of polygons being 
"cyclical" in nature.
\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(16,9)
	\put(-.7,6.5){\rotatebox{90}{First}}\put(-.4,6.4){\linethickness{1pt}\vector(0,-1){2}}
	\put(.2,8.2){Last}\put(1.5,8.4){\linethickness{1pt}\vector(1,0){2}}
	\Cfl{\Cg}\Cfd{\Cz}\Cfu{\Cg}
	\Cd[0,1]{6,1}{2.8}\Cd[0,1]{6,2}{1.8}\Cd[0,1]{6,3}{0.8}
	\Cd[1,0]{3,4}{2.8}\Cd[1,0]{4,4}{1.8}\Cd[1,0]{5,4}{0.8}
	\put(3.5,4.5){\line(3,-1){3}}
	\put(4.5,4.5){\line(2,-2){2}}
	\put(5.5,4.5){\line(1,-3){1}}
	%\Cd[1,-1]{5,2}{0.8}
\Cmlong\end{picture}\end{center}

\item Optimizations: we need to rotate that matrix to progress in the same direction as usual, that is towards bottom right.
{\color{red}}
\ole

@@

In practice, we don't go backward for larger values than 30, so we can replace $\min_{i<k<j}$ by $\min_{\max(i,j-30)<k<j}$ in the expressions of $VBI$.

>>>>

% ----------------------------------------------
\newpage
\subsubsection{Convex polygon triangulation}\ol
\item Problem: triangulating a polygon of $n$ vertices with least total cost for added edges. We denote the cost of adding an edge between the  pair of edges $i,j$ by $S(i,j)$, Where $S_{n \times n}$ is an upper triangular matrix with a 0 diagonal that is omitted, hence $|S|=\tfrac{n^2}{2}=N$.
\item Matrices: $M_{n\times n}, B_{n \times n}$ \textit{<<first vertex, last vertex>>} upper triangular including main diagonal. Note that if the vertex $n$ is the same as the vertex $0$.
\item Alphabets:\ul
	\item Input: $\Sigma(S_{(i,j)})=\{0..m\}$ with $m=\max_S(i,j) \forall i,j$ determined at runtime\footnote{We need to scan/have stats about $S$ and that's where LMS plays a role}.
	\item Cost matrix: $\Sigma(M)=\{0..z\}$ with $z = m \cdot (n-2)$ (we add at most $n-2$ edges).
	\item Backtrack matrix: $\Sigma(B)=\{stop, 0..n\}$ (the index of the edge we add)
	\ule
{\color{red}
\item Initialization: $M_{(i,i)}=0, M_{(i,i+1)}=0, B_{(i,i)}=stop \quad\forall i$

\item Recurrence: \[M_{(i,j)}=\left\{ S(i,j) + \max_{i<k<j}M_{(i,k)}+M_{(k,j)} \,\,\rule[-.75em]{.5pt}{2em}\,\,  k \right\} = B_{(i,j)} \]

@@

\item Visualization: the layout is the a matrix of size $n \times (2n-2)$, because of polygons being 
"cyclical" in nature.
\begin{center}\setlength{\unitlength}{.6cm}\begin{picture}(8,9)
	\put(-.7,6.5){\rotatebox{90}{First}}\put(-.4,6.4){\linethickness{1pt}\vector(0,-1){2}}
	\put(.2,8.2){Last}\put(1.5,8.4){\linethickness{1pt}\vector(1,0){2}}
	\Cfl{\Cg}
	\Cfd{\Cz}
	%\Cfu{\Cg}
	\Cd[0,1]{6,1}{2.8}\Cd[0,1]{6,2}{1.8}\Cd[0,1]{6,3}{0.8}
	\Cd[1,0]{3,4}{2.8}\Cd[1,0]{4,4}{1.8}\Cd[1,0]{5,4}{0.8}
	\put(3.5,4.5){\line(3,-1){3}}
	\put(4.5,4.5){\line(2,-2){2}}
	\put(5.5,4.5){\line(1,-3){1}}
	%\Cd[1,-1]{5,2}{0.8}
	\Cm %\Cmlong
\end{picture}\end{center}
}
\item Optimizations: we need the cost of edges between contiguous polygon vertices to have a cost of 0, so that we do not need to have special cases in the DP program (i.e. edges are <<already present>>, hence we cannot add them).
\ole

@@

With $\delta$ a lookup table. In practice, we don't go backward for larger values than 30, so we can replace $\min_{i<k<j}$ by $\min_{\max(i,j-30)<k<j}$ in the expressions of $VBI$.
